---
title: "Project 2"
format: html
editor: visual
---

## The Data

US Superstore date

```{r}
# Load libraries
library(tidyverse)
library(readxl)
```

```{r}
superstore <- read_excel("US Superstore data.xls", sheet = 1,
                         col_names = c("Row_ID", "Order_ID", "Order_Date",
                                       "Ship_Date", "Ship_Mode", "Customer_ID",
                                       "Customer_Name", "Segment", "Country",
                                       "City", "State", "Postal_Code",
                                       "Region", "Product_ID", "Category",
                                       "Sub_Category", "Product_Name", "Sales",
                                       "Quantity", "Discount", "Profit"),
                         col_types = c("numeric", "text", "date", "date",
                                       "text", "text", "text", "text", "text",
                                       "text", "text", "text", "text", "text",
                                       "text", "text", "text", "numeric",
                                       "numeric", "numeric", "numeric"),
                         skip = 1
)
superstore
```

Investigate the data

-   One-way and Two-way contingency tables

Row ID: Double

-   A unique ID for every entry, so it's not useful for telling us anything about the data.

```{r}
# One-Way Contingency Table
superstore |>
  group_by(Row_ID) |>
  summarize(count = n())
```

Order ID: Character

-   A unique ID for each order, however there's so many orders that this may not be very useful.

```{r}
# One-Way Contingency Table
superstore |>
  group_by(Order_ID) |>
  summarize(count = n())
```

Order Date and Ship Date: both Datetime

-   When the order was made and when it was shipped.

```{r}
# One-Way Contingency Table
superstore |>
  group_by(Order_Date) |>
  summarize(count = n())
```

```{r}
# One-Way Contingency Table
superstore |>
  group_by(Ship_Date) |>
  summarize(count = n())
```

Ship Mode: Character

-   The type of shipping, good for categorical grouping.

```{r}
# One-Way Contingency Table
superstore |>
  group_by(Ship_Mode) |>
  summarize(count = n())
```

Customer ID and Customer Name: both Character

-   A unique ID (and the name) for each customer, if there's a lot of customers this may or may not be useful.

```{r}
# One-Way Contingency Table
superstore |>
  group_by(Customer_Name) |>
  summarize(count = n())
```

Segment: Character

-   Type of customer, good for categorical grouping.

```{r}
# One-Way Contingency Table
superstore |>
  group_by(Segment) |>
  summarize(count = n())
```

Country: Character

-   I checked and there's only the United States

```{r}
# One-Way Contingency Table
superstore |>
  group_by(Country) |>
  summarize(count = n())
```

City: Character

-   City name, may be too many cities to be useful.

```{r}
# One-Way Contingency Table
superstore |>
  group_by(City) |>
  summarize(count = n())
```

State: Character

-   State name, could be a categorical grouping variable.

```{r}
# One-Way Contingency Table
superstore |>
  group_by(State) |>
  summarize(count = n())
```

Postal Code: Character?

-   The postal code for that area, probably even more numerous than City.

```{r}
# One-Way Contingency Table
superstore |>
  group_by(Postal_Code) |>
  summarize(count = n())
```

Region: Character

-   The region of the country of that area, good for categorical grouping.

```{r}
# One-Way Contingency Table
superstore |>
  group_by(Region) |>
  summarize(count = n())
```

Product ID and Product Name: Character

-   A unique ID (and name) for each product, too many of them to be useful.

```{r}
# One-Way Contingency Table
superstore |>
  group_by(Product_Name) |>
  summarize(count = n())
```

Category: Character

-   The category of the product, good for categorical grouping.

```{r}
# One-Way Contingency Table
superstore |>
  group_by(Category) |>
  summarize(count = n())
```

Sub-Category: Character

-   The subcategory of each product, could also use as a categorical grouping variable.

```{r}
# One-Way Contingency Table
superstore |>
  group_by(Sub_Category) |>
  summarize(count = n())
```

Sales: Double

-   The sales for that given purchase.

Quantity: Double

-   The number of that item bought in the purchase.

Discount: Double

-   The discount for that purchase.

Profit: Double

-   The profit made on that purchase.

<!-- -->

-   Numerical summaries (means, medians, sds, etc) for quantitative variables at levels of categorical variables.

-   Create at least six plots

    -   At least four of these should display multivariate information via the type of graph, by utilizing coloring, grouping, etc.

    -   All plots should have nice labels and titles.

    -   Some kind of faceting should be used somewhere.

    -   At least one plot should be a plot that we didn't cover in class.
